Fast and scalable Gaussian process modeling with applications to astronomical time series \vspace{-3\baselineskip}

### 2017-03-14 paper/ms.tex
Adding yet more references...

To demonstrate the applicability of this model, we apply it to infer the
asteroseismic parameters of the giant star KIC~11615890 observed by the
\kepler\ Mission.
The goal of this example is to show that, even for a low signal-to-noise
dataset with a short baseline, it is possible to infer asteroseismic
parameters with formal uncertainties that are consistent with the parameters
inferred with a much larger dataset.
Looking forward to \tess \citep{Campante2016}, we will infer $\nu_\mathrm{max}$ and $\Delta\nu$
using only one month of \kepler\ data and compare our results to the results
inferred from the full 4 year baseline of the mission.
For KIC~11615890, the published asteroseismic parameters based on the full
dataset are \citep{Pinsonneault:2014}
\begin{eqnarray}
    \nu_\mathrm{max} = 171.94 \pm 3.62 \,\mu\mathrm{Hz} \quad\mathrm{and}\quad
    \Delta\nu = 13.28 \pm 0.29 \,\mu\mathrm{Hz} \quad.
\end{eqnarray}
We randomly select a month-long segment of \kepler\ data, initialize our
\celerite\ model using a grid search in the parameter space, and then use
\emcee\ \citep{Foreman-Mackey:2013} to sample the joint posterior density for
the full set of parameters.
\Figure{astero-corner} shows the marginalized density for $\nu_\mathrm{max}$
and $\Delta\nu$ compared to the results from the literature.
\dfmtodo{Add some more discussion here.}

@@

Although we have in mind application of this fast method to stellar variability,
the method is general for one-dimensional GP problems, and may be applied to
other problems.    Within astrophysics, correlated noise (due to the environment,
detector, or modeling uncertainty) may be present in
gravitational wave time series, and so Gaussian processes may be a way to address
this problem \citep{Moore:2016}.  Accreting black holes show time series which
may be modeled by correlated noise \citep{Kelly:2014};  indeed, this was the
motivation for the original technique developed by Rybicki \& Press \citep{Rybicki:1992,Rybicki:1995}.
This approach may be broadly used for characterizing quasar variability
\citep{MacLeod:2010}, measuring time lags with reverberation mapping
\citep{Zu:2011}, and modeling time delays in multiply-imaged
gravitationally-lensed systems \citep{Press:1998}.

@@

There are three primary applications we envision for the \celerite\ formalism: 1).\ modeling
the variability of an astrophysical system to infer it properties; 2).\
accounting for astrophysical variability as a source of noise when trying to detect
additional phenomena; 3).\ interpolating or extrapolating variability to future or missing
times.  Some examples of the first are determining the characteristic
variability timescale of a quasar, measuring the asteroseismic variation of a star,
or detection of quasi-periodic variability in a high-energy source.  Examples of the
second are detecting transiting exoplanets and correcting for stellar activity in
radial velocity measurements. Examples of the third are measuring time delays in
multiply-imaged gravitationally lensed sources and reverberation mapping of AGN.

Our background is in studying transiting exoplanets, a field which has only recently begun
to adopt full covariance matrices in analyzing the noise in transiting planet lightcurves.
One promising early attempt at accounting for correlated noise was the wavelet approach
of \citet{Carter2009}.  Their technique allowed for a power-law spectrum for the
noise scaling as $\omega^{-\alpha}$, which unfortunately is not a normalizable, nor
physically-accurate, description of stellar variability.  Their approach runs in
${\cal O}(N)$ time for $\alpha=1$, which is referred to as `flicker', $1/r$, or pink 
noise;  however, this particular type of noise overpredicts the power spectrum at low 
frequency compared to the Solar granulation power spectrum.  It also cannot describe 
quasi-periodic noise.

@@

>>>>

To demonstrate the applicability of this model, we apply it to infer the
asteroseismic parameters of the giant star KIC~11615890 observed by the
\kepler\ Mission.
The goal of this example is to show that, even for a low signal-to-noise
dataset with a short baseline, it is possible to infer asteroseismic
parameters with formal uncertainties that are consistent with the parameters
inferred with a much larger dataset.
Looking forward to \tess \citep{Campante:2016}, we will infer $\nu_\mathrm{max}$ and $\Delta\nu$
using only one month of \kepler\ data and compare our results to the results
inferred from the full 4 year baseline of the mission.
For KIC~11615890, the published asteroseismic parameters based on the full
dataset are \citep{Pinsonneault:2014}
\begin{eqnarray}
    \nu_\mathrm{max} = 171.94 \pm 3.62 \,\mu\mathrm{Hz} \quad\mathrm{and}\quad
    \Delta\nu = 13.28 \pm 0.29 \,\mu\mathrm{Hz} \quad.
\end{eqnarray}
We randomly select a month-long segment of \kepler\ data, initialize our
\celerite\ model using a grid search in the parameter space, and then use
\emcee\ \citep{Foreman-Mackey:2013} to sample the joint posterior density for
the full set of parameters.
\Figure{astero-corner} shows the marginalized density for $\nu_\mathrm{max}$
and $\Delta\nu$ compared to the results from the literature.
\dfmtodo{Add some more discussion here.}

@@

Although we have in mind application of this fast method to stellar variability,
the method is general for one-dimensional GP problems, and may be applied to
other problems.  Accreting black holes show time series which
may be modeled by correlated noise \citep{Kelly:2014};  indeed, this was the
motivation for the original technique developed by Rybicki \& Press \citep{Rybicki:1992,Rybicki:1995}.
This approach may be broadly used for characterizing quasar variability
\citep{MacLeod:2010}, measuring time lags with reverberation mapping
\citep{Zu:2011}, and modeling time delays in multiply-imaged
gravitationally-lensed systems \citep{Press:1998}.

@@

There are three primary applications we envision for the \celerite\ formalism: 1).\ modeling
the variability of an astrophysical system to infer it properties; 2).\
accounting for astrophysical variability as a source of noise when trying to detect
additional phenomena; 3).\ interpolating or extrapolating variability to future or missing
times.  Some examples of the first are determining the characteristic
variability timescale of a quasar \citep{Kelly:2014}, measuring the asteroseismic variation of a star
\citep{Corsaro:2014}, measurement of granulation noise \citep{Bastien:2013,Kallinger:2016},
detection of quasi-periodic variability in a high-energy source \citep{McAllister:2016},
or classification of variable objects \citep{Zinn:2016}.  Examples of the
second are detecting transiting exoplanets and correcting for stellar activity in
radial velocity measurements \citep{Haywood:2014,Rajpaul:2015}. Examples of the third are 
measuring time delays in
multiply-imaged gravitationally lensed sources \citep{Tewes:2013} and reverberation mapping of AGN.

Our background is in studying transiting exoplanets, a field which has only recently begun
to adopt full covariance matrices in analyzing the noise in transiting planet lightcurves.
One promising early attempt at accounting for correlated noise was the wavelet approach
of \citet{Carter:2009}.  Their technique allowed for a power-law spectrum for the
noise scaling as $\omega^{-\alpha}$, which unfortunately is not a normalizable, nor
physically-accurate, description of stellar variability.  Their approach runs in
${\cal O}(N)$ time for $\alpha=1$, which is referred to as `flicker', $1/r$, or pink 
noise;  however, this particular type of noise overpredicts the power spectrum at low 
frequency compared to the Solar granulation power spectrum.  It also cannot describe 
quasi-periodic noise.

@@

In the future we hope to expand the \celerite\ technique to multi-dimensional
datasets.  The formalism presented could, for example, be used to model
the covariance between multiple measured quantities, such as in regression of photometric
time series versus astrometric precision when correcting for systematic errors
due to detector sensitivity variations \citep{Aigrain:2016}, in correcting
radial velocity datasets for stellar activity \citep{Haywood:2014,Rajpaul:2015},
or in detection of gravitational waves in the presence of systematic errors
\citep{Moore:2016}.
